\Large \bf One-Step, Three-Factor Authentication With Custom-Fit, In-Ear EEG

### 2017-02-15 paper.tex
Trying again with small fixes and sample size discussion

\subsection{Passthought authentication}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor \cite{Chuang2013b}. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrrrr}
& \multicolumn{6}{c}{Impedances [k\(\Omega\)]} \\
\cline{2-7}
& \multicolumn{3}{|c|}{Left ear} & \multicolumn{3}{c|}{Right ear} \\
P & Concha & Front & Back & Concha & Front & Back \\
\hline
2 & 34.1 & 10.2 & 12.8 & 27.8 & 16.0 & 16.3\\
3 & 21.1 & 20.9 & 19.0 & 13.5 & 11.3 & 19.5\\
4 & 14.1 & 11.9 & 9.7 & 11.0 & 11.1 & 13.3\\
5 & 17.2 & 21.9 & 10.3 & 32.6 & 12.5 & 11.6\\
6 & 18.7 & 10.0 & 8.4 & 14.8 & 11.5 & 8.9\\
7 & 91.5 & \textgreater1000 & 21.5 & 33.5 & 26.4 & 31.0\\
\end{tabular}
\end{center}
\caption{Electrical impedances measured while P1 wore each other participant's custom-fitted earpieces.}
\label{tab:imposter_impedances}
\end{table}

@@

Our findings demonstrate the apparent feasibility of single earpiece, achieving good results with only three electrodes and a reference, all on the left ear. FARs and FRRs are low across all participants and tasks, with FARs overall lower than FRRs. Participants' best-performing passthoughts typically seeing no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

@@

In our analysis, some notable patterns emerged. First, \textit{breathe} tended to be the best-performing task among participants. Classifiers overall distinguished the breath task even compared to breath tasks from other participants, implying that the task is expressed differently for each participant, i.e. that this task has an inherence factor sufficient for authentication, even though the task does not have a knowledge factor. Second, we were able to achieve good results by generating feature vectors based on only 500ms (300 voltage readings across the three electrodes). This short timespan is somewhat surprising, given that some tasks (like songs) presumably rely on changes or patterns over a longer period of time.

Furthermore, the use of conductive gel results in low impedances on other participants' custom-fit earpieces, despite the uniquness of ear canal shapes between individuals \cite{Akkermans2005}.
Nevertheless, classifiers appear to resists spoofing attacks, indicating that task-related signals are unique to individuals.
In the future, the custom-fit earbuds could include a hardware keypair to sign authentication attempts, providing a more secure method for establishing the posession factor in our three-factor, one-step scheme.

>>>>

\subsection{Passthoughts Authentication}
The use of EEG as a biometric signal for user authentication has a short history.
In 2005, Thorpe et al. motivate and outline the design of a passthoughts system \cite{Thorpe2005}.
Since 2002, a number of independent groups have achieved 99-
100\% authentication accuracy using multi-channel sensors placed on the scalp \cite{Poulos2002,Marcel2007a,Palaniappan2008,Ashby2011}.
In 2013, one group showed that 99\% authentication accuracy can also be
achieved using a consumer-grade single-channel sensor \cite{Chuang2013b}. In particular, the
lack of signal diversity from multiple EEG channels can be overcome by allowing
the users to choose their own personalized passthoughts (e.g., sing their favorite
song in their head). There are two significant consequences of this result. First,
the passthoughts approach is no longer constrained by the high cost (\textgreater \$10,000 USD)
and low usability (gel-based electrodes; aesthetic challenges of an EEG cap) of
medical-grade multi-channel devices. Second, because users can choose and
easily change their secret mental task, this approach can support one-step two-
factor authentication via the simultaneous presentation of the inherence factor
(brainwave signatures due to the unique folding structures of the cortex) and the
knowledge factor (the secret mental task) \cite{Chuang2014}.

@@

\begin{table}[h]
\begin{center}
\begin{tabular}{lrrrrrr}
& \multicolumn{6}{c}{Impedances [k\(\Omega\)]} \\
\cline{2-7}
& \multicolumn{3}{|c|}{\textbf{Left ear}} & \multicolumn{3}{c|}{\textbf{Right ear}} \\
\textbf{P} & \textbf{Concha} & \textbf{Front} & \textbf{Back} & \textbf{Concha} & \textbf{Front} & \textbf{Back} \\
\hline
2 & 34.1 & 10.2 & 12.8 & 27.8 & 16.0 & 16.3\\
3 & 21.1 & 20.9 & 19.0 & 13.5 & 11.3 & 19.5\\
4 & 14.1 & 11.9 & 9.7 & 11.0 & 11.1 & 13.3\\
5 & 17.2 & 21.9 & 10.3 & 32.6 & 12.5 & 11.6\\
6 & 18.7 & 10.0 & 8.4 & 14.8 & 11.5 & 8.9\\
7 & 91.5 & \textgreater1000 & 21.5 & 33.5 & 26.4 & 31.0\\
\end{tabular}
\end{center}
\caption{Electrical impedances measured while P1 wore each other participant's custom-fitted earpieces.}
\label{tab:imposter_impedances}
\end{table}

@@

Our findings demonstrate the apparent feasibility of single earpiece, achieving good results with only three electrodes and a reference, all on or around the left ear. FARs and FRRs are low across all participants and tasks, with FARs overall lower than FRRs, a desirable pattern in terms of authenticating access to potentially sensitive information. Participants' best-performing passthoughts typically see no errors in our training. Furthermore, no spoofed attacks were successful in our cursory analysis.

This study was conducted with a small, relatively homogeneous sample of participants. We recognize this weakness and agree that more participants from diverse sample would greatly improve our understanding and interpretation of the capabilities of this technology. This does not lessen the importance of the work reported here however, as this sample size is comparable to that of previous scalp-based EEG passthoughts work\cite{Ashby2011, Marcel2007a, Palaniappan2008, palaniappan2006electroencephalogram, Poulos2002, Chuang2013b}, the only other in-ear EEG passthoughts study we are aware of \cite{curranpassthoughts} (notably a generic fit system was used here), and other custom-fit in-ear EEG research \cite{kidmose2013study}. The fitting and manufacturing of custom-fit earpieces was the main limitation to increasing our sample size, and may very well pose a limitation in adoption of such a technology. Recently, however there have been developments in at-home kits for creating one's own custom-fitted earpieces \cite{voix2015settable} that could help overcome this barrier.

@@

In our analysis, some notable patterns emerged. First, \textit{breathe} tended to be the best-performing task among participants. Classifiers overall distinguished the breathe task even compared to breathe tasks from other participants, implying that the task is expressed differently for each participant, i.e. that this task has an inherence factor sufficient for authentication, even though the task does not have a knowledge factor. Second, we were able to achieve good results by generating feature vectors based on only 500ms (300 voltage readings across the three electrodes). This short timespan is somewhat surprising, given that some tasks (like songs) presumably rely on changes or patterns over a longer period of time.

Furthermore, the use of conductive gel results in low impedances on other participants' custom-fit earpieces, despite the uniqueness of ear canal shapes between individuals \cite{Akkermans2005}.
Nevertheless, classifiers appear to resists spoofing attacks, indicating that task-related signals are unique to individuals.
In the future, the custom-fit earbuds could easily include a hardware keypair to sign authentication attempts, providing a more secure method for establishing the possession factor in our three-factor, one-step scheme.
